### 2018-02-12 submission/manuscript.Rmd
update discussion and conclusion

***Minimal Bray-Curtis Differences are Detected by Cycle Number:***
Overall, there was little difference between each respective 5-cycle increment at higher cycle numbers, for both fecal and mock samples, and this was consistent across the different sub-samplings [Figure 3]. There were only two differences between 5-cycle increments that were identified. First, there were differences for the same fecal sample between 20 versus 25 cycles that was independent of HiFi DNA polymerase but dependent on sub-sampling depth (sub-sampled to 1000 = `r round(median(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`) (median (25% - 75% quantile)), sub-sampled to 5000 = `r round(median(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`), sub-sampled to 10000 = `r round(median(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`)) [Figure 3A]. Second, for the mocks, where data is available, there were larger difference between 20 and 25 cycles (sub-sampled to 1000 = `r round(median(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`)) [Figure 3B]. However, these differences between the next 5-cycle increment do not persist once 25 cycles are reached [Figure 3].

Using PERMANOVA to test for differences within HiFi DNA polymerase groups based on cycle number, only Phusion had cycle dependent differences at 1000 and 5000 sub-sampling depth (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)` and `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 5000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`, respectively). Phusion was one of only two HiFi DNA polymerases that managed to have samples for the 1000 sub-sampling depth at 15 cycles. Next, we assessed whether there were any major differences between 5-cycle increments within each sample. We found that there was no detectable difference in Bray-Curtis index when comparing to the previous 5-cycle increment (P-value `r ifelse(min(bray_sim_to_prev_cycle_kruskal_results$bh, na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05). However, Phusion at 1000 sub-sampling depth had a P-value = `r format(filter(bray_sim_to_prev_cycle_kruskal_results, sub_sample_level == 1000, taq == "PHU")[, "pvalue"], digits = 1, nsmall = 1)` before multiple comparison correction. It should be noted that at higher sub-sampling depths these differences in Bray-Curtis indices disappear for both differences in cycle number and within 5-cycle increments within an individual.

@@

***Chimeric Sequences are HiFi DNA Polymerase Dependent and Correlate with Number of OTUs:***
After chimera removal using VSEARCH and removal of sequences that did not classify as bacteria, we assessed the percentage of sequences that were still chimeric within our mock community. At all levels of sub-sampling and cycle number there were significant differences between the HiFi DNA polymerase used (P-value `r ifelse(max(filter(kruskal_chimera_prev, sub_sample_level > 500)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S9]. Differences between Platinum and all other HiFi DNA polymerases accounted for the vast majority of these differences independent of cycle number and sub-sampling depth when using a Dunn's post-hoc test [Table S10]. Generally, across sub-sampling depth and cycle number Accuprime^TM^ had the lowest chimera prevalence of all the HiFi DNA Polymerases regardless of whether 'pre.cluster' or chimera removal with VSEARCH had been used [Figure 5].

For all DNA polymerase, a positive correlation was observed between chimeric sequences and number of OTUs, with this correlation being strongest for Accuprime, Platinum and Phusion HiFi DNA Polymerase [Figure 6]. The R^2^ value between the number of OTUs and chimeric sequences did not change from the use of 'pre.cluster' and the removal of chimeras with VSEARCH [Figure 6]. Taken together, this data suggests that a strong correlation exists between the number of OTUs and the prevalence of chimeric sequences.

@@

Our observations build upon previous studies [@taq_Gohl_2016] by showing that different HiFi DNA polymerases have both significant differences in the number of OTUs and that changes in total OTUs correlate with chimeras not removed after sequence processing [Figure 1-2 & 5]. This is important since many diversity metrics rely on the total number of OTUs as part of their calculations and changes to the total number of OTUs could drastically change the results and conclusions. Our observations show that HiFi DNA polymerase can have a noticeable affect on the OTUs generated and these differences are consistent across sub-sampling depth and PCR cycle number [Figure 2-4]. These differences were observed in high biomass samples, where biases introduced by such components like kit contamination may have less of an effect, suggesting that these differences may be exacerbated in low biomass samples.

Although we did not observe strong differences in the Bray-Curtis index the data suggests that there may be differences between 15 and 20 cycles versus higher cycle numbers, such as 30x, that are commonly used. There was few differences within individuals between corresponding 5-cycle increments (e.g. 15 to 20, 20 to 25, etc.). However, there is a clear trend that suggests that 20 cycles is very different then 25 cycles [Figure 3]. We may just not have enough samples to adequately assess the true magnitude difference between 20 and 25 cycle communities and higher cycle numbers analyzed. These findings, in conjunction with the PERMANOVA results, suggest that cycle number can change bacterial community calculations but that these differences are minimal once 25 cycles are reached. For some DNA polymerase, increasing the sub-sampling depth may reduce some of these observed community differences at lower cycle numbers.

Increasing the cycle number also exacerbated chimera prevalence differences between the different HiFi DNA polymerases [Figure 5]. The chimera prevalence was strongly correlated with the number of OTUs which is relied upon heavily for many community metric calculations. However, Bray-Curtis analysis with PERMANOVA showed few differences based on DNA polymerase. It is possible that many of the increased number of OTUs are not highly abundant allowing the Bray-Curtis index to be able to successfully down-weight these respective OTUs [@bc_index_Minchin1987]. The choice of downstream diversity metric could be an important consideration in helping to mitigate these observed changes due to high chimera prevalence in specific HiFi DNA polymerases (e.g. Platinum).

Our observations suggest that there are clear HiFi DNA polymerase dependent differences in both per base error rate and chimeras that cannot be removed using bioinformatic approaches [Figure 4 & 5]. Although it may be a natural assumption that the variation may be due to the DNA polymerase family, the highest chimera rate, from Platinum, was a family A polymerase while the lowest, from Accuprime, was also an A family polymerase [@polymerase_Ishino_2014]. In fact, from the material safety data sheet (MSDS), the differences between the two mixes is not immediately apparent. Both Accuprime and Platinum contain a recombinant *Taq* DNA polymerase, a *Pycrococcus* spp GB-D polymerase and a platinum *Taq* antibody. It is possible that differences in how the recombinant *Taq* was generated could be the main reason for the differences in chimera rate.

@@

Our findings show that measures that rely on the number of OTUs will be specific for a particular study due to both HiFi DNA polymerase and cycle number chosen. Due in part to these consderations findings may not be easily generalized to other studies investigating a similar area. Care should be taken when choosing a HiFi DNA polymerase for 16S rRNA gene surveys since intrinsic differences can change the number of OTUs observed as well as potentially influence diversity based metrics that do not down-weight rare observations.

>>>>

***Minimal Bray-Curtis Differences are Detected by Cycle Number:***
For both fecal and mock samples, there was little difference between each respective 5-cycle increment at higher cycle numbers and this was consistent across the different sub-samplings [Figure 3]. Overall, there were only two differences between 5-cycle increments that were identified. First, there were differences for the same fecal sample between 20 versus 25 cycles that was independent of HiFi DNA polymerase but dependent on sub-sampling depth (sub-sampled to 1000 = `r round(median(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`) (median (25% - 75% quantile)), sub-sampled to 5000 = `r round(median(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`), sub-sampled to 10000 = `r round(median(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`)) [Figure 3A]. Second, for the mocks, where data is available, there were larger difference between 20 and 25 cycles (sub-sampled to 1000 = `r round(median(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`)) [Figure 3B]. However, these differences between the next 5-cycle increment do not persist once 25 cycles are reached [Figure 3].

Using PERMANOVA to test for differences within HiFi DNA polymerase groups based on cycle number, only Phusion had cycle dependent differences at 1000 and 5000 sub-sampling depth (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)` and `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 5000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`, respectively). Phusion was also one of only two HiFi DNA polymerases that managed to have samples for the 1000 sub-sampling depth at 15 cycles. Next, we assessed whether there were any major differences between 5-cycle increments within each sample. We found that there was no detectable difference in Bray-Curtis index when comparing to the previous 5-cycle increment (P-value `r ifelse(min(bray_sim_to_prev_cycle_kruskal_results$bh, na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05). However, Phusion at 1000 sub-sampling depth had a P-value = `r format(filter(bray_sim_to_prev_cycle_kruskal_results, sub_sample_level == 1000, taq == "PHU")[, "pvalue"], digits = 1, nsmall = 1)` before multiple comparison correction. It should be noted that at higher sub-sampling depths these differences in Bray-Curtis indices disappear for both differences in cycle number and within 5-cycle increments within an individual.

@@

***Chimeric Sequences are HiFi DNA Polymerase Dependent and Correlate with Number of OTUs:***
After chimera removal using VSEARCH and removal of sequences that did not classify as bacteria, we assessed the percentage of sequences that were still chimeric within our mock community. At all levels of sub-sampling and cycle number there were significant differences between the HiFi DNA polymerase used (P-value `r ifelse(max(filter(kruskal_chimera_prev, sub_sample_level > 500)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S9]. Differences between Platinum and all other HiFi DNA polymerases accounted for the vast majority of these differences independent of cycle number and sub-sampling depth when using a Dunn's post-hoc test [Table S10]. Across sub-sampling depth and cycle number Accuprime^TM^ had the lowest chimera prevalence of all the HiFi DNA Polymerases regardless of whether 'pre.cluster' or chimera removal with VSEARCH had been used [Figure 5].

For all HiFi DNA polymerases, a positive correlation was observed between chimeric sequences and number of OTUs, with this correlation being strongest for Accuprime, Platinum and Phusion HiFi DNA Polymerase [Figure 6]. The R^2^ value between the number of OTUs and chimeric sequences did not change from the use of 'pre.cluster' and the removal of chimeras with VSEARCH [Figure 6]. Taken together, this data suggests that a strong correlation exists between the number of OTUs and the prevalence of chimeric sequences.

@@

Our observations build upon previous studies [@taq_Gohl_2016] by showing that different HiFi DNA polymerases have both significant differences in the number of OTUs and that changes in total OTUs correlate with chimeras not removed after sequence processing [Figure 1-2 & 5]. HiFi DNA polymerase dependent differences in total number of OTUs are important to consider since many diversity metrics rely on this metric as part of their calculations. Our observations show that HiFi DNA polymerase can have a noticeable affect on the OTUs generated and these differences are consistent across sub-sampling depth and PCR cycle number [Figure 2-4]. These differences were observed in high biomass samples, where biases introduced by such components like kit contamination may have less of an effect, suggesting that these differences may be exacerbated in low biomass samples.

Although we did not observe strong differences in the Bray-Curtis index the data suggests that there may be differences between 15 and 20 cycles versus higher cycle numbers, such as 30x, that are commonly used. There was few differences within individuals between corresponding 5-cycle increments (e.g. 15 to 20, 20 to 25, etc.). However, there is a clear trend that suggests that 20 cycles is very different then 25 cycles [Figure 3]. We may just not have enough samples to adequately assess the true magnitude difference between 20 and 25 cycle communities and higher cycle numbers analyzed. These findings, in conjunction with the PERMANOVA results, suggest that cycle number can change bacterial community calculations but that these differences are minimal once 25 cycles are reached. For some HiFI DNA polymerases, increasing the sub-sampling depth may reduce some of these observed community differences at lower cycle numbers.

Increasing the cycle number also exacerbated chimera prevalence differences between the different HiFi DNA polymerases [Figure 5]. The chimera prevalence was strongly correlated with the number of OTUs which is relied upon heavily for many community metric calculations. However, Bray-Curtis analysis with PERMANOVA showed few differences based on HiFi DNA polymerase. It is possible that many of the increased number of OTUs are not highly abundant allowing the Bray-Curtis index to be able to successfully down-weight these respective OTUs [@bc_index_Minchin1987]. The choice of downstream diversity metric could be an important consideration in helping to mitigate these observed changes due to high chimera prevalence in specific HiFi DNA polymerases (e.g. Platinum).

Our observations suggest that there are clear HiFi DNA polymerase dependent differences in both per base error rate and chimera prevalence that cannot be removed using bioinformatic approaches [Figure 4 & 5]. Although it may be a natural assumption that the variation may be due to the DNA polymerase family, the highest chimera rate, from Platinum, was a family A polymerase while the lowest, from Accuprime, was also an A family polymerase [@polymerase_Ishino_2014]. In fact, from the material safety data sheet (MSDS), the differences between the two mixes is not immediately apparent. Both Accuprime and Platinum contain a recombinant *Taq* DNA polymerase, a *Pycrococcus* spp GB-D polymerase and a platinum *Taq* antibody. It is possible that differences in how the recombinant *Taq* was generated could be the main reason for the differences in chimera rate.

@@

Our findings show that measures that rely on the number of OTUs will be dependent on both HiFi DNA polymerase and cycle number chosen. Due in part to these consderations, studies investigating similar areas may not be easily generalized with each other. Care should be taken when choosing a HiFi DNA polymerase for 16S rRNA gene surveys since intrinsic differences can change the number of OTUs observed and influence diversity based metrics that do not down-weight rare observations.

### 2018-02-12 submission/manuscript.tex
update discussion and conclusion

\textbf{\emph{Minimal Bray-Curtis Differences are Detected by Cycle
Number:}} Overall, there was little difference between each respective
5-cycle increment at higher cycle numbers, for both fecal and mock
samples, and this was consistent across the different sub-samplings
{[}Figure 3{]}. There were only two differences between 5-cycle
increments that were identified. First, there were differences for the
same fecal sample between 20 versus 25 cycles that was independent of
HiFi DNA polymerase but dependent on sub-sampling depth (sub-sampled to
1000 = 0.51 (0.4 - 0.79) (median (25\% - 75\% quantile)), sub-sampled to
5000 = 0.43 (0.33 - 0.63), sub-sampled to 10000 = 0.4 (0.24 - 0.43))
{[}Figure 3A{]}. Second, for the mocks, where data is available, there
were larger difference between 20 and 25 cycles (sub-sampled to 1000 =
0.88 (0.42 - 0.91)) {[}Figure 3B{]}. However, these differences between
the next 5-cycle increment do not persist once 25 cycles are reached
{[}Figure 3{]}.

Using PERMANOVA to test for differences within HiFi DNA polymerase
groups based on cycle number, only Phusion had cycle dependent
differences at 1000 and 5000 sub-sampling depth (P-value = 0.03 and
0.01, respectively). Phusion was one of only two HiFi DNA polymerases
that managed to have samples for the 1000 sub-sampling depth at 15
cycles. Next, we assessed whether there were any major differences
between 5-cycle increments within each sample. We found that there was
no detectable difference in Bray-Curtis index when comparing to the
previous 5-cycle increment (P-value \textgreater{} 0.05). However,
Phusion at 1000 sub-sampling depth had a P-value = 0.02 before multiple
comparison correction. It should be noted that at higher sub-sampling
depths these differences in Bray-Curtis indices disappear for both
differences in cycle number and within 5-cycle increments within an
individual.

@@

\textbf{\emph{Chimeric Sequences are HiFi DNA Polymerase Dependent and
Correlate with Number of OTUs:}} After chimera removal using VSEARCH and
removal of sequences that did not classify as bacteria, we assessed the
percentage of sequences that were still chimeric within our mock
community. At all levels of sub-sampling and cycle number there were
significant differences between the HiFi DNA polymerase used (P-value
\textless{} 0.05) {[}Table S9{]}. Differences between Platinum and all
other HiFi DNA polymerases accounted for the vast majority of these
differences independent of cycle number and sub-sampling depth when
using a Dunn's post-hoc test {[}Table S10{]}. Generally, across
sub-sampling depth and cycle number Accuprime\textsuperscript{TM} had
the lowest chimera prevalence of all the HiFi DNA Polymerases regardless
of whether `pre.cluster' or chimera removal with VSEARCH had been used
{[}Figure 5{]}.

For all DNA polymerase, a positive correlation was observed between
chimeric sequences and number of OTUs, with this correlation being
strongest for Accuprime, Platinum and Phusion HiFi DNA Polymerase
{[}Figure 6{]}. The R\textsuperscript{2} value between the number of
OTUs and chimeric sequences did not change from the use of `pre.cluster'
and the removal of chimeras with VSEARCH {[}Figure 6{]}. Taken together,
this data suggests that a strong correlation exists between the number
of OTUs and the prevalence of chimeric sequences.

@@

Our observations build upon previous studies (Gohl et al., 2016) by
showing that different HiFi DNA polymerases have both significant
differences in the number of OTUs and that changes in total OTUs
correlate with chimeras not removed after sequence processing {[}Figure
1-2 \& 5{]}. This is important since many diversity metrics rely on the
total number of OTUs as part of their calculations and changes to the
total number of OTUs could drastically change the results and
conclusions. Our observations show that HiFi DNA polymerase can have a
noticeable affect on the OTUs generated and these differences are
consistent across sub-sampling depth and PCR cycle number {[}Figure
2-4{]}. These differences were observed in high biomass samples, where
biases introduced by such components like kit contamination may have
less of an effect, suggesting that these differences may be exacerbated
in low biomass samples.

Although we did not observe strong differences in the Bray-Curtis index
the data suggests that there may be differences between 15 and 20 cycles
versus higher cycle numbers, such as 30x, that are commonly used. There
was few differences within individuals between corresponding 5-cycle
increments (e.g.~15 to 20, 20 to 25, etc.). However, there is a clear
trend that suggests that 20 cycles is very different then 25 cycles
{[}Figure 3{]}. We may just not have enough samples to adequately assess
the true magnitude difference between 20 and 25 cycle communities and
higher cycle numbers analyzed. These findings, in conjunction with the
PERMANOVA results, suggest that cycle number can change bacterial
community calculations but that these differences are minimal once 25
cycles are reached. For some DNA polymerase, increasing the sub-sampling
depth may reduce some of these observed community differences at lower
cycle numbers.

Increasing the cycle number also exacerbated chimera prevalence
differences between the different HiFi DNA polymerases {[}Figure 5{]}.
The chimera prevalence was strongly correlated with the number of OTUs
which is relied upon heavily for many community metric calculations.
However, Bray-Curtis analysis with PERMANOVA showed few differences
based on DNA polymerase. It is possible that many of the increased
number of OTUs are not highly abundant allowing the Bray-Curtis index to
be able to successfully down-weight these respective OTUs (Minchin,
1987). The choice of downstream diversity metric could be an important
consideration in helping to mitigate these observed changes due to high
chimera prevalence in specific HiFi DNA polymerases (e.g.~Platinum).

Our observations suggest that there are clear HiFi DNA polymerase
dependent differences in both per base error rate and chimeras that
cannot be removed using bioinformatic approaches {[}Figure 4 \& 5{]}.
Although it may be a natural assumption that the variation may be due to
the DNA polymerase family, the highest chimera rate, from Platinum, was
a family A polymerase while the lowest, from Accuprime, was also an A
family polymerase (Ishino \& Ishino, 2014). In fact, from the material
safety data sheet (MSDS), the differences between the two mixes is not
immediately apparent. Both Accuprime and Platinum contain a recombinant
\emph{Taq} DNA polymerase, a \emph{Pycrococcus} spp GB-D polymerase and
a platinum \emph{Taq} antibody. It is possible that differences in how
the recombinant \emph{Taq} was generated could be the main reason for
the differences in chimera rate.

>>>>

\textbf{\emph{Minimal Bray-Curtis Differences are Detected by Cycle
Number:}} For both fecal and mock samples, there was little difference
between each respective 5-cycle increment at higher cycle numbers and
this was consistent across the different sub-samplings {[}Figure 3{]}.
Overall, there were only two differences between 5-cycle increments that
were identified. First, there were differences for the same fecal sample
between 20 versus 25 cycles that was independent of HiFi DNA polymerase
but dependent on sub-sampling depth (sub-sampled to 1000 = 0.51 (0.4 -
0.79) (median (25\% - 75\% quantile)), sub-sampled to 5000 = 0.43 (0.33
- 0.63), sub-sampled to 10000 = 0.4 (0.24 - 0.43)) {[}Figure 3A{]}.
Second, for the mocks, where data is available, there were larger
difference between 20 and 25 cycles (sub-sampled to 1000 = 0.88 (0.42 -
0.91)) {[}Figure 3B{]}. However, these differences between the next
5-cycle increment do not persist once 25 cycles are reached {[}Figure
3{]}.

Using PERMANOVA to test for differences within HiFi DNA polymerase
groups based on cycle number, only Phusion had cycle dependent
differences at 1000 and 5000 sub-sampling depth (P-value = 0.03 and
0.01, respectively). Phusion was also one of only two HiFi DNA
polymerases that managed to have samples for the 1000 sub-sampling depth
at 15 cycles. Next, we assessed whether there were any major differences
between 5-cycle increments within each sample. We found that there was
no detectable difference in Bray-Curtis index when comparing to the
previous 5-cycle increment (P-value \textgreater{} 0.05). However,
Phusion at 1000 sub-sampling depth had a P-value = 0.02 before multiple
comparison correction. It should be noted that at higher sub-sampling
depths these differences in Bray-Curtis indices disappear for both
differences in cycle number and within 5-cycle increments within an
individual.

@@

\textbf{\emph{Chimeric Sequences are HiFi DNA Polymerase Dependent and
Correlate with Number of OTUs:}} After chimera removal using VSEARCH and
removal of sequences that did not classify as bacteria, we assessed the
percentage of sequences that were still chimeric within our mock
community. At all levels of sub-sampling and cycle number there were
significant differences between the HiFi DNA polymerase used (P-value
\textless{} 0.05) {[}Table S9{]}. Differences between Platinum and all
other HiFi DNA polymerases accounted for the vast majority of these
differences independent of cycle number and sub-sampling depth when
using a Dunn's post-hoc test {[}Table S10{]}. Across sub-sampling depth
and cycle number Accuprime\textsuperscript{TM} had the lowest chimera
prevalence of all the HiFi DNA Polymerases regardless of whether
`pre.cluster' or chimera removal with VSEARCH had been used {[}Figure
5{]}.

For all HiFi DNA polymerases, a positive correlation was observed
between chimeric sequences and number of OTUs, with this correlation
being strongest for Accuprime, Platinum and Phusion HiFi DNA Polymerase
{[}Figure 6{]}. The R\textsuperscript{2} value between the number of
OTUs and chimeric sequences did not change from the use of `pre.cluster'
and the removal of chimeras with VSEARCH {[}Figure 6{]}. Taken together,
this data suggests that a strong correlation exists between the number
of OTUs and the prevalence of chimeric sequences.

@@

Our observations build upon previous studies (Gohl et al., 2016) by
showing that different HiFi DNA polymerases have both significant
differences in the number of OTUs and that changes in total OTUs
correlate with chimeras not removed after sequence processing {[}Figure
1-2 \& 5{]}. HiFi DNA polymerase dependent differences in total number
of OTUs are important to consider since many diversity metrics rely on
this metric as part of their calculations. Our observations show that
HiFi DNA polymerase can have a noticeable affect on the OTUs generated
and these differences are consistent across sub-sampling depth and PCR
cycle number {[}Figure 2-4{]}. These differences were observed in high
biomass samples, where biases introduced by such components like kit
contamination may have less of an effect, suggesting that these
differences may be exacerbated in low biomass samples.

Although we did not observe strong differences in the Bray-Curtis index
the data suggests that there may be differences between 15 and 20 cycles
versus higher cycle numbers, such as 30x, that are commonly used. There
was few differences within individuals between corresponding 5-cycle
increments (e.g.~15 to 20, 20 to 25, etc.). However, there is a clear
trend that suggests that 20 cycles is very different then 25 cycles
{[}Figure 3{]}. We may just not have enough samples to adequately assess
the true magnitude difference between 20 and 25 cycle communities and
higher cycle numbers analyzed. These findings, in conjunction with the
PERMANOVA results, suggest that cycle number can change bacterial
community calculations but that these differences are minimal once 25
cycles are reached. For some HiFI DNA polymerases, increasing the
sub-sampling depth may reduce some of these observed community
differences at lower cycle numbers.

Increasing the cycle number also exacerbated chimera prevalence
differences between the different HiFi DNA polymerases {[}Figure 5{]}.
The chimera prevalence was strongly correlated with the number of OTUs
which is relied upon heavily for many community metric calculations.
However, Bray-Curtis analysis with PERMANOVA showed few differences
based on HiFi DNA polymerase. It is possible that many of the increased
number of OTUs are not highly abundant allowing the Bray-Curtis index to
be able to successfully down-weight these respective OTUs (Minchin,
1987). The choice of downstream diversity metric could be an important
consideration in helping to mitigate these observed changes due to high
chimera prevalence in specific HiFi DNA polymerases (e.g.~Platinum).

Our observations suggest that there are clear HiFi DNA polymerase
dependent differences in both per base error rate and chimera prevalence
that cannot be removed using bioinformatic approaches {[}Figure 4 \&
5{]}. Although it may be a natural assumption that the variation may be
due to the DNA polymerase family, the highest chimera rate, from
Platinum, was a family A polymerase while the lowest, from Accuprime,
was also an A family polymerase (Ishino \& Ishino, 2014). In fact, from
the material safety data sheet (MSDS), the differences between the two
mixes is not immediately apparent. Both Accuprime and Platinum contain a
recombinant \emph{Taq} DNA polymerase, a \emph{Pycrococcus} spp GB-D
polymerase and a platinum \emph{Taq} antibody. It is possible that
differences in how the recombinant \emph{Taq} was generated could be the
main reason for the differences in chimera rate.
